Hysteresis in human computation:\\ how one task affects another

### 2014-12-03 docs/publications/2014.09_SCIENCE/man.tex
another proof-read of the manuscript

Faced with the scale and complexity of the data revolution, researchers and
industry are increasingly relying on \textit{human computation}.  Microtask
platforms, like Amazon Mechanical Turk, act like a human compute server,
making it possible to access human cognition quickly and inexpensively. 
This has lead to broad adoption across a spectrum of disciplines
in the humanities and computer science, in research into medical imaging
technologies, and in many applications where expert opinion would normally be 
sought.  Here we show that an inherent
feature of microtask platforms, the repetition of similar tasks,
acts as a strong, but as yet unexplored source of bias.  
These \textit{intertask effects}, rival the
severity of overtly framing a task's purpose.  
Using image labeling as a canonical microtask, we found that prior 
tasks influences the vocabulary that a worker uses when labeling 
subsequent images in profound an nuanced ways.  Prior tasks not only 
alter what a worker focuses on, 
but also modulate the richness and specialization of her vocabulary.  
While they can be a source of systematic bias, our
results suggest that intertask effects can be leveraged
to hone worker focus and specificity.  Though we focused on microtask 
platforms due to their emerging prominence, our results apply to  
citizen science, crowdsourcing, and any application involving repetitive 
tasks relying on human judgment.
\end{sciabstract}

@@

Unlike CPUs and GPUs, the experiments we present here suggest HPUs
are subject to \textit{hysteresis}, meaning that their output depends
on the history of their inputs.  Given the
increasing prevalence of microtask-based methodologies, other factors 
affecting HPU outputs have received considerable attention.
This includes various task-design parameters such as the level of 
pay \cite{kazai2013analysis}, the training \cite{le2010ensuring} and 
pre-screening \cite{paolacci2010running} of workers, and user-interface 
design factors \cite{Finnerty2013}.  
Researchers have also investigated \textit{framing} effects, 
such as the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or altering the problem description
\cite{thibodeau2013natural}.  Results in psychology show that people are 
susceptible to various kinds of priming 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, 
task-repetition effects \cite{Gass1999549,sohn2001task}.  But to our 
knowledge, no study has investigated the practical ramifications 
of intertask effects on human computation.

We investigated intertask effects using image-labeling tasks on the Amazon 
Mechanical Turk (MTurk) microtask platform.  Image labeling tasks are among
the most common MTurk tasks \cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, a trend which will probably continue 
in the long term because of their role in computer vision research 
\cite{5543192}.
In our experiments, workers performed a set of \textit{initial tasks}, 
followed by a set of \textit{test tasks}. Crucially, no distinction was
made between these sets from the perspective of the worker. We varied the 
initial tasks (for example, having workers label images of food or objects), 
and observed the ensuing effects on responses in test tasks.  
As a point of 
comparison, we also included treatments in which we framed the tasks in terms 
of different purposes.

The results showed that intertask effects are stronger than framing effects.  
Framing effects were comparable to intertask effects only when we had the
workers actively reiterate the frame by means of a combo-box input.
Using the wordnet corpus (\#), we found that intertask 
effects can hone a worker's focus, and boost the specialization
and lexical richness of her labels.  Thus, while intertask effects may be a 
serious source of bias, they can also be leveraged to help achieve
expert-level performance in crowdsourcing when properly controlled.

\paragraph{Experimental setup.}
\begin{figure}
	\centering
	\includegraphics[scale=0.7]{figs/images.pdf}
	\caption{
		Examples of images used for various experiments and treatments:
		A) initial tasks for \textit{task1:food}, B) initial tasks for 
		\textit{task1:obj}, C) test tasks for treatments in \textit{exp1}, 
		D) initial tasks for \textit{task2:food}, 
		E) initial tasks for \textit{task2:cult},  
		F) test tasks for \textit{exp2}.  The full set of experimental 
		materials is shown in the supplementary material.
	}

@@

We performed two experiments, \textit{exp1} and \textit{exp2}, which consisted
of various pairs of treatments.  The treatments in each pair were subjected to
the same exposure modality, either initial tasks or framing, but differed in
the content of the exposure.  Table \ref{table:treatments}  
summarizes the treatments and introduces a naming convention that we use 
in what follows.
In \textit{exp1}, workers in the treatment pair \textit{task1} initially
labeled 5 images containing either food (treatment \textit{task1:food}), or 
objects, (treatment \textit{task1:obj}), and then performed test tasks, in 
which they 
labeled images containing both food and objects (see Fig.~\ref{fig:task} for 
examples of the images used). In \textit{exp1}, we included two other 
treatment pairs, \textit{frame1} and \textit{echo1}, in which workers did not 
perform initial tasks, but instead were given information which framed the 
purpose of the tasks.  Workers in \textit{frame1} were told that the tasks 
were ``Funded by the laboratory for the visual perception of 
\{ Food and Ingredients $\vert$ Objects and Tools~\}''.  In \textit{echo1},
we sharpened the frame, by indicating that ``The purpose of this study is to 
understand the visual perception of 
\{ Food and Ingredients $\vert$ Objects and Tools~\}'' and then asking the
worker to reiterate the purpose of the task using a combo-box input.

In contrast to \textit{exp1}, which was based on concepts of food and
objects, \textit{exp2} was based food and \textit{culture}.
Since culture is more abstract, one might expect different effects 
in \textit{exp2}.  We again compared
intertask effects (in treatment pair \textit{task2}) to framing effects 
(\textit{frame2}), but with a slight variation in the framing treatments.
In \textit{frame2}, workers were not only exposed to a frame, but also 
performed a set of initial tasks before the test tasks.  Since its purpose 
was still
to investigate framing effects, the initial tasks in \textit{frame2}
were not varied. Thus, \textit{frame2} provides insight into the moderating 
effect that intervening tasks might have on framing effects.

In all treatments, workers were shown images one at a time, and provided
5 labels for each.  From the perspective of the worker, there was no
distinction or interruption between the initial and test tasks.  We used
workers' unique identifiers to ensure that they only participated in a 
one treatment of a single experiment.

@@

\paragraph*{Measuring computational hysteresis.}
While our study builds on research into the psychological mechanisms
of task performance, there is a crucial difference in the question at issue. 
Our goal is to measure the practical ramifications of intertask effects on 
the microtask methodology. Whereas a statistically significant effect, even 
if small, might be sufficient to distinguish one psychological mechanism 
from another (\#), in our study 
\textit{effect size} is central.  Thus our main goal is not merely to reject 
the null hypothesis that initial tasks have no effect
(we do this in the supplementary text).

To couch intertask effects in a practical yet general setting, we imagine
that, after a worker performs a task, her response is used to make some 
kind of yes-or-no decision, $\mathcal{D}$.  For example, this decision 
could be whether or not to flag a certain comment
below a news article as inappropriate; but for now, we leave 
$\mathcal{D}$ unspecified.  Since worker's responses may
vary, repeating the same task might lead to different outcomes for 
$\mathcal{D}$.  We shall define the notion of \textit{bias} in terms of the
effect that prior tasks have on the probability of a positive outcome
from $\mathcal{D}$.

The nature of $\mathcal{D}$ will affect our measurement of bias.
If $\mathcal{D}$ has random output, or if it always outputs ``yes'' 
irrespective of the worker's response, then of course we will 
find zero bias.  Likewise, there is an upper limit to the bias induced on $\mathcal{D}$. In a precise sense, the bias of $\mathcal{D}$ can only be as
great as the bias inherent in the worker's response 
(this is formalized in the supplementary text).  Therefore, to estimate the 
bias induced in workers' outputs, we adopt an 
adversarial approach: we seek to construct the decision that 
is the most susceptible to workers' biases. In the supplementary text, we 
show that this
problem can be formulated as a  machine learning problem, wherein one
constructs a classifier that infers the exposure of a worker from her 
responses in test tasks.  The bias, $\theta$, induced by 
prior exposures, can then be bounded using the classifier's accuracy, 
$\eta$ (\#):
\begin{equation}
	\theta \geq 2\eta - 1
	\label{l1}
\end{equation}

\paragraph{The strength and dynamics of intertask effects.}
For each treatment pair, we trained a naive Bayes classifier (\# in the 
supplementary material we discuss the choice of classifier, the
approach to measuring accuracy, and draw a comparison to results using 
a support vector machine in place of the naive Bayes classifier) to 
distinguish workers from one treatment
(e.g \textit{task1:food}) from those of the conjugate treatment 
(e.g. \textit{task1:obj}).  In \textit{task1} and \textit{task2}, intertask 
effects profoundly biased worker outputs, whereas we could not detect 
a statistically significant bias induced by framing in \textit{frame1} and 
\textit{frame2} (see Fig.~\ref{fig:theta}).  The bias induced
by framing was only of a comparable level to that induced by intertask 
effects in \textit{echo1}.  
Moreover, asking workers to reiterate the
frame, as was done in \textit{echo1}, is probably interpreted as
a direct signal from the requester that the frame should be acted upon.  In that sense, 
\textit{echo1} is perhaps more indicative of the effects of an 
\textit{instruction}.  This shows that
intertask effects profoundly affect the responses of workers, producing an
effect that is stronger than framing, and on par with the combination of 
framing and reiteration.

\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/theta.pdf}
	\caption{
		Lower-bound estimates for the  worst-case bias introduced in image
		labeling tasks due to worker exposure effects.  The estimate was 
		based on the performance of a multinomial naive Bayes classifier,
		trained to infer HPU exposures. A) Bias induced through various
		modalities and prime concepts (see Table 1). B) Bias induced in 
		\textit{task1}, as a function of test task position.  Error bars
		show the one-tailed 95\% confidence interval.
	}
	\label{fig:theta}
\end{figure}

To observe the dynamics of intertask effects, we trained a naive Bayes
classifier to distinguish workers in \textit{task1} based on responses to
individual test tasks.  We permuted 
the test tasks to eliminate effects arising from differences in the images 
therein (more details are provided in the supplementary text). The first test 
task shows the greatest bias, after which the effects drop off, but 
substantial bias persists through the fifth test task
(see Fig. \ref{fig:theta}B).  Presumably, this decay profile reflects the
fact that workers are also influenced by their exposure to the test tasks, 
which washes out the effects of initial tasks.

@@

First, we used wordnet to identify nouns depicting foods (\# We 
operationalized food-references by taking all hyponyms of the synsets 
\texttt{food.n.01} and \texttt{food.n.02}, with the addition of ethnic foods 
learned by crawling a recipe website, as described in the supplementary text),
to see whether 
workers from the various food-exposed treatments (\textit{*:food}), 
provided more references to food than their conjugate treatments, overall.
The \textit{exp1:*:food} treatments
did produce more food references, but, unexpectedly, 
the number of food references in \textit{task2:food} was
less than that in \textit{task2:cult}. Also, the effects on the
number of food references was, overall, somewhat smaller than one might 
have expected given the extent of bias observed for \textit{task1}, 
\textit{echo1}, and \textit{task2}.  This means that task and frame exposures
induce other effects which are more subtle than a bulk shift in number 
of references to the primed concept (food).

@@

\paragraph{Positive and negative priming.}
Taken together, the results of the experiments presented here can be 
explained through a combination of positive and negative priming.
Positive priming (usually simply ``priming'') occurs when a prior stimulus 
predisposes a person to give certain responses in an ensuing task, and
is often observed as an increase in the speed or accuracy of a response, or
the ability to recognize objects in a briefer, quieter, or noisier signal (\#).
Negative priming occurs when repeated exposure to a stimulus that is 
perceived to be non-salient begins to be ignored (\#).

Workers exposed to images containing food will be (positively) primed, 
as memories, concepts, and vocabulary related to food are activated.  
However, as the worker labels successive images containing food, the basic 
fact that an image contains food does not appear to be salient, since it 
does not serve to distinguish one image from another.  Thus, the most 
generic reference to the fact that food appears, namely the label ``food'', 
is suppressed, as by negative priming, while more specialized
references are activated.  Whether workers in \textit{*:food} treatments
produce an excess of food references overall depends in part on the balance 
of these factors.  The reduction in food references from workers in 
\textit{task2:food} relative
to those in \textit{task2:cult} may also be related to the novelty that 
workers in 
\textit{task2:cult} experience when they first encounter food in the test 
tasks.

@@

Based on our observations throughout, we submit the following 
recommendations for the consideration of those in the position of designing 
any microtask, citizen science, or crowdsourcing initiative:
\begin{enumerate}
	\item{
		Require active feedback to emphasize important details, 
		as we did in \textit{echo1}.  Similarly, use training tasks rather
		than (only) examples.
	}
	\item{
		Although ``clean'' examples may be useful to provide clarity,
		training tasks should contain the same kinds of noise or otherwise
		distracting features as are expected in the real tasks.
		This exposure will negatively prime the worker against non-salient 
		features.
	}
	\item{
		\textit{Validation tasks}, for which the correct output is known,
		are often used as a quality control measure.
		These should be dually considered as \textit{calibration tasks}.  
		Feedback, for both correct and incorrect responses may further 
		reinforce the desired competency. \label{item:valid}
	}
	\item{
		A special case of item~\ref{item:valid} applies when workers are 
		required to detect rare occurrences.
		For example, in the detection of pre-ictal (pre-seizure) EEG traces, 
		most traces are not negative (\#).  In the absence of positive 
		examples, the target concept may degrade or drift.
		Drawing validation/calibration tasks from the underrepresented class,
		while providing feedback might help sustain the target concept.
	}
	\item{
		For more precise classifications, it may help to begin with
		an initial coarse classification.  For example, in detecting 
		inappropriate comments, workers might initially choose from  
		\texttt{okay}, \texttt{inappropriate}, and \texttt{borderline}.  
		Following the coarse classification, the 
		\texttt{borderline} examples could be batched together.  Workers could
		then calibrate to finer distinctions while coding a more homogeneous 
		set of images.
	}
\end{enumerate}

@@

\paragraph{Conclusion.}
We have shown that intertask effects are stronger than framing effects, and
as strong as the effects induced when framing is combined with reiteration.
Intertask and framing effects act on workers' vocabulary in subtle ways.  
Both increase the richness of workers' vocabulary in reference a target
concept, with a more pronounced effect induced by task exposure.
Both exposures also increase the specialization of workers' vocabulary in 
reference to the target concept.  Framing, and other aspects of task design 
have received substantial methodological scrutiny; our results show that 
intertask effects, which were overlooked until now, merit similar 
consideration.  Although we found that intertask effects can be a major 
source of bias, they might be used to drive tune workers' attention and 
specificity relative to
a target concept.  Microtask work, citizen science, and other crowdsourcing
initiatives, as well task requiring expert opinions, often share the common 
feature that they rely on repetitive human judgments.  Therefore, such 
applications all stand to benefit from the findings and suggestions 
presented here.

@@

\subsection*{S1: Experimental materials}
The tasks for both experiments were presented as a series of slides.  In
both cases, the first slide consisted of a brief set of instructions, followed
by the frame if shown (only some treatments involved a frame), see 
Fig. S\ref{fig:hit_preamble}.  Following the instructions and prime, a set
of initial tasks was shown (except for the treatments in \textit{frame2}
and \textit{echo}) followed by (in all cases) a set of test tasks.  In
\textit{exp1}, two kinds of frame were used, one simple frame, and one where
the worker was required to echo the frame using a combo-box input.  The
instructions, both kinds of frame, and an example task slide, are shown 
in Fig. S1.
\begin{figure}
	\includegraphics[scale=0.8]{figs/tasks.pdf}
	\label{fig:hit_preamble}
	\caption{Examples of A) instructions; B) a frame, as shown in 
		\textit{frame1:obj}, and similar to those shown in 
		\textit{frame1:food} and \textit{frame2:food}, and 
		\textit{frame2:cult};  
		C) an echoed frame, as used in \textit{echo:obj}, and similar to that
		used in \textit{echo:food}; and D) an 
		example of an image labeling task.
	}
\end{figure}

@@

	For the purpose of training and testing a naive Bayes classifier, we 
	performed an additional preprocessing step.  We tagged words with the
	position in which they had been entered (i.e. which of the five text 
	inputs in the task interface) as well as the test tasks in which the word 
	had been provided.
	So, if the word ``wine'' was entered into the second text-input for 
	the third test-task, after preprocessing, the feature ``3\_2\_wine'' would
	appear.  Prepending the task number onto words was simply a means to 
	retain correct attribution, which itself was essential since providing a
	word during one task is not equivalent to providing the same word during 
	another task.

@@

\paragraph{Selection of the naive Bayes classifier.}
A great deal of work has been done in the application of classifiers to
text documents, which provides the background for our choice of classifier.
The Support Vector Machine (SVM) approach is known to perform well, better in 
general than naive Bayes.  However, three considerations drove us to use
a naive Bayes classifier in favor of SVM.

First, we used 119 examples (workers) per treatment, but the number of 
features
for each worker (i.e. the number of distinct labels, or modifications 
thereof produced in preprocessing), was generally in the thousands.  
This tends to
be problematic for many classifier algorithms, including SVM, because it
leads to very high variance models, which manifests as overfitting and poor
performance.  To deal with this, some approach to culling features is 
generally used, but this raises the issue of how to decide the culling in 
a principled manner, generally adding further hyperparameters which could
be optimized.  However, one of the strengths of the naive Bayes classifier
is that it's performance is not generally degraded from the inclusion of 
large numbers of features, even when the number of examples is comparatively
small.  We expected this feature of the naive Bayes classifier to be 
advantageous in our case.

Second, the standard implementation of the naive Bayes classifier has no 
hyperparameters to optimize.  This allows us to forgo the common practice 
of holding out a test set, and instead use cross-validation as a method to
estimate the true classification error, \textit{since we did not perform
any kind of model selection or optimization whatsoever using our data}.
This point is crucial, because if we did optimize our classifier during 
cross-validation, then the validation error would no longer provide an 
unbiased estimate of the true classification error, but would instead lead
to us over-reporting exposure effects.

>>>>

Faced with the scale and complexity of the data revolution, researchers and
industry are increasingly relying on \textit{human computation}.  Microtask
platforms, like Amazon Mechanical Turk, act like a human compute server,
making it possible to access human cognition quickly and inexpensively. 
This has lead to broad adoption across a spectrum of disciplines
in the humanities and computer science, as well as in research into medical 
imaging technologies and other applications where expert opinion would 
normally be sought.  Here we show that an inherent
feature of microtask platforms, the repetition of similar tasks,
acts as a strong, but as yet unexplored source of bias.  
These \textit{intertask effects}, rival the
severity of overtly framing a task's purpose.  
Using image-labeling as a canonical microtask, we found that prior 
tasks influences the vocabulary that a worker uses when labeling 
subsequent images in profound an nuanced ways.  Prior tasks not only 
alter what a worker focuses on, 
but also modulate the richness and specialization of her vocabulary.  
While intertask effects can be a source of systematic bias, our
results suggest that, when properly controlled, they can be leveraged
to hone worker focus and specificity.  Though we focused on microtask 
platforms due to their emerging prominence, our results apply to  
citizen science, crowdsourcing, and any application involving repetitive 
tasks relying on human judgment.
\end{sciabstract}

@@

Unlike CPUs and GPUs, the experiments we present here suggest HPUs
are subject to \textit{hysteresis}, meaning that their outputs depend
on the history of their inputs.  Given the
increasing prevalence of microtask-based research methods, other factors 
affecting HPU outputs have received considerable attention.
This includes various task-design parameters such as the level of 
pay \cite{kazai2013analysis}, the training \cite{le2010ensuring} and 
pre-screening \cite{paolacci2010running} of workers, and user-interface 
design factors \cite{Finnerty2013}.  
Researchers have also investigated \textit{framing} effects, 
such as the effects of describing the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, or altering the problem description
\cite{thibodeau2013natural}.  Results in psychology show that people are 
susceptible to various kinds of priming 
\cite{BJOP1796,No2007,beller1971priming}, and, in particular, 
task-repetition effects \cite{Gass1999549,sohn2001task}.  But to our 
knowledge, no study has investigated the practical ramifications 
of intertask effects on human computation.

We investigated intertask effects using image-labeling tasks on the Amazon 
Mechanical Turk (MTurk) microtask platform.  Image-labeling tasks are among
the most common MTurk tasks \cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}, a trend which will probably continue 
in the long term because of their role in computer vision research 
\cite{5543192}.
In our experiments, workers performed a set of \textit{initial tasks}, 
followed by a set of \textit{test tasks}. Crucially, no distinction was
made between these sets from the perspective of the worker. We varied the 
initial tasks (for example, having workers label images of food or objects), 
and observed the ensuing effects on responses in test tasks.  
As a point of 
comparison, we also included treatments in which we framed the tasks in terms 
of different purposes.

The results showed that intertask effects are stronger than framing effects.  
Framing effects were comparable to intertask effects only when the workers 
were asked to actively reiterate the frame by means of a combo-box input.
Using the wordnet corpus (\#), we found that intertask 
effects can hone a worker's focus, and boost the specialization
and lexical richness of her labels.  Thus, while intertask effects may be a 
serious source of bias, they can also be leveraged to help achieve
expert-level performance in crowdsourcing, when properly controlled.

\paragraph{Experimental setup.}
\begin{figure}
	\centering
	\includegraphics[scale=0.7]{figs/images.pdf}
	\caption{
		Examples of images used for various experiments and treatments:
		A) initial tasks for \textit{task1:food}, B) initial tasks for 
		\textit{task1:obj}, C) test tasks for treatments in \textit{exp1}, 
		D) initial tasks for \textit{task2:food}, 
		E) initial tasks for \textit{task2:cult},  
		F) test tasks for \textit{exp2}.  The full set of experimental 
		materials is shown in the supplementary text.
	}

@@

We performed two experiments, \textit{exp1} and \textit{exp2}, which consisted
of various pairs of treatments.  The treatments in each pair were subjected to
the same exposure modality, either initial tasks or framing, but differed in
the content of the exposure.  Table \ref{table:treatments}  
summarizes the treatments and introduces a naming convention that we use 
in what follows.
In \textit{exp1}, workers in the treatment pair \textit{task1} initially
labeled 5 images containing either food (treatment \textit{task1:food}), or 
objects (\textit{task1:obj}), and then performed test tasks, in 
which they 
labeled images containing both food and objects (see Fig.~\ref{fig:task} for 
examples of the images used). In \textit{exp1}, we included two other 
treatment pairs, \textit{frame1} and \textit{echo1}, in which workers did not 
perform initial tasks, but instead were given information which framed the 
purpose of the tasks.  Workers in \textit{frame1} were told that the tasks 
were ``Funded by the laboratory for the visual perception of 
\{ Food and Ingredients $\vert$ Objects and Tools~\}''.  In \textit{echo1},
we sharpened the frame, by indicating that ``The purpose of this study is to 
understand the visual perception of 
\{ Food and Ingredients $\vert$ Objects and Tools~\}'' and then asking the
worker to reiterate the purpose of the task using a combo-box input.

In contrast to \textit{exp1}, which was based on concepts of food and
objects, \textit{exp2} was based food and \textit{culture}.
Since culture is more abstract, one might expect different effects 
in \textit{exp2}.  In \textit{exp2}, we again investigated both
intertask effects (in treatment pair \textit{task2}) and framing effects 
(\textit{frame2}), but with a slight variation in the framing treatments.
In \textit{frame2}, workers were not only exposed to a frame, but also 
performed a set of initial tasks before the test tasks.  Since its purpose 
was still
to investigate framing effects, the initial tasks in \textit{frame2}
were not varied. Thus, \textit{frame2} provides insight into the moderating 
effect that intervening tasks might have on framing effects.

In all treatments, workers were shown images one at a time, and had to
provide 5 labels for each.  From the perspective of the worker, there was no
distinction or interruption between the initial and test tasks.  We used
workers' unique identifiers to ensure that they only participated in a 
one treatment of a single experiment.

@@

\paragraph*{Measuring computational hysteresis.}
While our study builds on research into the psychological mechanisms
of task performance, there is a crucial difference in our present goal. 
Our goal is to measure the practical ramifications of intertask effects on 
the microtask methodology. Whereas a statistically significant effect, even 
if small, might be sufficient to distinguish one psychological mechanism 
from another (\#), in our study, 
\textit{effect size} is central.  Thus our main goal is not merely to reject 
the null hypothesis that initial tasks have no effect
(we do this in the supplementary text).

To couch intertask effects in a practical yet general setting, we imagine
that, after a worker performs a task, her response is used to make some 
kind of yes-or-no decision, $\mathcal{D}$.  For example, this decision 
could be whether or not to flag a certain comment
below a news article as inappropriate; but for now, we leave 
$\mathcal{D}$ unspecified.  Since workers' responses may
vary, repeating the same task might lead to different outcomes for 
$\mathcal{D}$.  This decision not only models the reality that responses are
actually used---introducing $\mathcal{D}$ allows us to define of \textit{bias}
in a practical yet general way. We shall define bias in terms of the
change in the probability of a positive outcome from $\mathcal{D}$, induced 
by prior exposures.

The kind of decision that $\mathcal{D}$ represents will affect our 
measurement of bias.  For instance, if $\mathcal{D}$ is random, or if 
it always generates ``yes'' 
irrespective of the worker's response, then of course we will 
find zero bias.  Likewise, there is an upper limit to the bias induced on $\mathcal{D}$. In a precise sense, the bias of $\mathcal{D}$ can only be as
great as the bias inherent in the worker's response 
(this is formalized in the supplementary text).  Therefore, to estimate the 
bias induced in workers' outputs, we adopt an 
adversarial approach: we seek to construct the decision that 
is the most susceptible to workers' biases. In the supplementary text, we 
show that this can be formulated as a  machine learning problem, wherein one
constructs a classifier that infers the prior exposure of a worker from her 
responses in test tasks.  The bias, $\theta$, induced by 
prior exposures, can then be bounded using the classifier's accuracy, 
$\eta$ (\#):
\begin{equation}
	\theta \geq 2\eta - 1
	\label{l1}
\end{equation}

\paragraph{The strength and dynamics of intertask effects.}
For each treatment pair, we trained a na\"ive Bayes classifier (\# in the 
supplementary material we discuss the choice of classifier, the
approach to measuring accuracy, and draw a comparison to the results obtained 
if a support vector machine is used.) to distinguish workers from one 
treatment (e.g \textit{task1:food}) from those of the conjugate treatment 
(e.g. \textit{task1:obj}).  In \textit{task1} and \textit{task2}, intertask 
effects profoundly biased worker outputs, by approximately 30 and 50\% 
respectively, whereas we could not detect 
a statistically significant bias induced by framing in \textit{frame1} and 
\textit{frame2} (see Fig.~\ref{fig:theta}A).  The bias induced
by framing was only appreciable in \textit{echo1}, reaching approximately 
35\%.  
We suspect that asking workers to reiterate the
frame, as was done in \textit{echo1}, might be interpreted by workers as
a direct signal that the frame should be acted upon.  In that sense, 
\textit{echo1} is perhaps more indicative of the effects of an 
\textit{instruction}.  These observations show that
prior tasks profoundly influence the responses of workers, producing an
effect that is stronger than framing, and on par with the combination of 
framing and reiteration.

\begin{figure}
	\centering
	\includegraphics[scale=1]{figs/theta.pdf}
	\caption{
		Lower-bound estimates for the  worst-case bias introduced in image
		labeling tasks due to worker exposure effects.  The estimate was 
		based on the performance of a multinomial na\"ive Bayes classifier,
		trained to infer HPU exposures. A) Bias induced through various
		modalities and prime concepts (see Table 1). B) Bias induced in 
		\textit{task1}, as a function of test task position.  Error bars
		show the one-tailed 95\% confidence interval.
	}
	\label{fig:theta}
\end{figure}

To observe the dynamics of intertask effects, we trained a na\"ive Bayes
classifier to distinguish workers in \textit{task1} based on responses to
individual test tasks.  We permuted 
the test tasks to eliminate effects arising from differences in the images 
therein (more details are provided in the supplementary text). The first test 
task shows the greatest bias, after which the effects drop somewhat, but 
substantial bias persists through the fifth test task
(see Fig. \ref{fig:theta}B).  Presumably, this decay profile reflects the
fact that workers are also influenced by their exposure to the test tasks, 
which begins to wash out the effects of initial tasks.

@@

First, we used wordnet to identify nouns depicting foods (\# We 
operationalized food-references by taking all hyponyms of the synsets 
\texttt{food.n.01} and \texttt{food.n.02}, with the addition of ethnic foods 
learned by crawling a recipe website, as described in the supplementary text),
to see whether 
workers from the various food-exposed treatments (\textit{*:food}), 
provided more references to food than their conjugate treatments, overall.
The \textit{exp1:*:food} treatments
did produce more food references, but, unexpectedly, 
the number of food references in \textit{task2:food} was
less than that in \textit{task2:cult} (see Fig.~\ref{fig:specificity}A). Also, the effects on the
number of food references was, overall, somewhat smaller than one might 
expect given the extent of bias observed in Fig.~\ref{fig:theta}A.  This means that task and frame exposures
induce other effects which are more subtle than a bulk shift in the number 
of references to the primed concept (food).

@@

\paragraph{Positive and negative priming.}
Taken together, the results of the experiments presented here can be 
explained through a combination of positive and negative priming.
Positive priming (usually simply ``priming'') occurs when a prior stimulus 
predisposes a person to give certain responses in an ensuing task, and
is often observed as an increase in the speed or accuracy of a response, or
the ability to recognize objects in a briefer, quieter, or noisier signal (\#).
Negative priming occurs when, after being repeatedly exposed to a stimulus 
that is perceived to be non-salient, the person begins to ignore that 
stimulus (\#).

Workers exposed to images containing food will be (positively) primed, 
as memories, concepts, and vocabulary related to food are activated.  
However, we argue that as the worker labels successive images containing food,
the basic 
fact that an image contains food will not appear to be salient, since it 
does not serve to distinguish one image from another.  Thus, the most 
generic references to that fact, such as the label ``food'', 
are suppressed, as by negative priming, while more specialized
references are activated.  Whether workers in \textit{*:food} treatments
produce an excess of food references overall depends in part on the balance 
of these factors.  The reduction in food references from workers in 
\textit{task2:food} relative
to those in \textit{task2:cult} may also be related to the novelty that 
workers in 
\textit{task2:cult} experience when they first encounter food in the test 
tasks.

@@

\paragraph{Recommendations and future work.} 
Based on our observations throughout, we submit the following 
recommendations for the consideration of people designing 
any microtask, citizen science, or crowdsourcing initiative:
\begin{enumerate}
	\item{
		Require active feedback to emphasize important details about the 
		tasks, as we did in \textit{echo1}.  Similarly, use training tasks 
		rather than only examples.
	}
	\item{
		Although ``clean'' examples may be useful to provide clarity,
		use training tasks that contain the same kinds of noise or otherwise
		distracting features as are expected in the real tasks.
		This exposure will negatively prime the worker against non-salient 
		features.
	}
	\item{
		\textit{Validation tasks}, for which the correct output is known,
		are often used as a quality control measure.
		Treat these as \textit{calibration tasks}, and use them to regulate
		the focus of the worker.
		Feedback, for both correct and incorrect responses may further 
		reinforce the desired competency. \label{item:valid}
	}
	\item{
		A special case of item~\ref{item:valid} applies when workers are 
		required to detect rare occurrences.
		For example, in the detection of pre-ictal (pre-seizure) EEG traces, 
		most traces are negative (\#).  In the absence of positive 
		examples, the target concept may degrade or drift.
		Draw calibration tasks from the underrepresented class,
		and provide feedback, to help sustain the worker's target concept.
	}
	\item{
		For more precise classifications, it may help to begin with
		an initial coarse classification.  For example, in detecting 
		inappropriate comments, provide the choices   
		\texttt{okay}, \texttt{inappropriate}, and \texttt{borderline},
		and incentivise inter-worker agreement
		Next, batch the \texttt{borderline}.  Coding this more homogeneous
		borderline set will help workers to calibrate to the finer 
		distinctions necessary for the task.
	}
\end{enumerate}

@@

\paragraph{Conclusion.}
We have shown that intertask effects are stronger than framing effects, and
as strong as the effects induced when framing is combined with reiteration.
Intertask and framing effects act on workers' vocabulary in subtle ways.  
Both increase the richness of workers' vocabulary in reference a target
concept, with a more pronounced effect induced by task exposure.
Both exposures also increase the specialization of workers' vocabulary in 
reference to the target concept.  Framing, and other aspects of task design 
have received substantial methodological scrutiny; our results show that 
intertask effects, which were overlooked until now, merit similar 
attention.  Although we found that intertask effects can be a major 
source of bias, they might be used to tune workers' attention and 
specificity relative to
a target concept.  Microtask work, citizen science, and other crowdsourcing
initiatives, as well task requiring expert opinions, often rely on repetitive 
human judgments.  Therefore, such applications all stand to benefit from the 
findings and suggestions presented here.

@@

\subsection*{S1: Experimental materials}
The tasks for both experiments were presented as a series of slides.  In
both cases, the first slide consisted of a brief set of instructions, followed
by the frame if shown (only some treatments involved a frame), see 
Fig. S\ref{fig:hit_preamble}.  Following the instructions and prime, a set
of initial tasks was shown (except for the treatments in \textit{frame2}
and \textit{echo}) followed by (in all cases) a set of test tasks.  In
\textit{exp1}, two kinds of frame were used, one simple frame, and one where
the worker was required to echo the frame using a combo-box input.  The
instructions, both kinds of frame, and an example task slide, are shown 
in Fig. S1.
\begin{figure}
	\includegraphics[scale=0.8]{figs/tasks.pdf}
	\label{fig:hit_preamble}
	\caption{Examples of A) instructions; B) a frame, as shown in 
		\textit{frame1:obj}, and similar to those shown in 
		\textit{frame1:food} and \textit{frame2:food}, and 
		\textit{frame2:cult};  
		C) an echoed frame, as used in \textit{echo:obj}, and similar to that
		used in \textit{echo:food}; and D) an 
		example of an image-labeling task.
	}
\end{figure}

@@

	For the purpose of training and testing a na\"ive Bayes classifier, we 
	performed an additional preprocessing step.  We tagged words with the
	position in which they had been entered (i.e. which of the five text 
	inputs in the task interface) as well as the test tasks in which the word 
	had been provided.
	So, if the word ``wine'' was entered into the second text-input for 
	the third test-task, after preprocessing, the feature ``3\_2\_wine'' would
	appear.  Prepending the task number onto words was simply a means to 
	retain correct attribution, which itself was essential since providing a
	word during one task is not equivalent to providing the same word during 
	another task.

@@

\paragraph{Selection of the na\"ive Bayes classifier.}
A great deal of work has been done in the application of classifiers to
text documents, which provides the background for our choice of classifier.
The Support Vector Machine (SVM) approach is known to perform well, better in 
general than na\"ive Bayes.  However, three considerations drove us to use
a na\"ive Bayes classifier in favor of SVM.

First, we used 119 examples (workers) per treatment, but the number of 
features
for each worker (i.e. the number of distinct labels, or modifications 
thereof produced in preprocessing), was generally in the thousands.  
This tends to
be problematic for many classifier algorithms, including SVM, because it
leads to very high variance models, which manifests as overfitting and poor
performance.  To deal with this, some approach to culling features is 
generally used, but this raises the issue of how to decide the culling in 
a principled manner, generally adding further hyperparameters which could
be optimized.  However, one of the strengths of the na\"ive Bayes classifier
is that it's performance is not generally degraded from the inclusion of 
large numbers of features, even when the number of examples is comparatively
small.  We expected this feature of the na\"ive Bayes classifier to be 
advantageous in our case.

Second, the standard implementation of the na\"ive Bayes classifier has no 
hyperparameters to optimize.  This allows us to forgo the common practice 
of holding out a test set, and instead use cross-validation as a method to
estimate the true classification error, \textit{since we did not perform
any kind of model selection or optimization whatsoever using our data}.
This point is crucial, because if we did optimize our classifier during 
cross-validation, then the validation error would no longer provide an 
unbiased estimate of the true classification error, but would instead lead
to us over-reporting exposure effects.
